A Gravitationally Lensed Supernova with an Observable Two-Decade Time Delay

### 2021-03-28 Paper/ResponseToReferees/response_coverletter.tex

\documentclass[11pt]{article}
%\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{soul}
%\usepackage{graphicx}
%\usepackage{lineno}
%\linenumbers

%\usepackage{hyperref}
%\usepackage{booktabs}

%\usepackage{amssymb}
\usepackage{xspace}
%\usepackage{graphicx}
%\usepackage{todonotes}

%\usepackage{geometry}

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%\definechangesauthor[name=1, color=blue]{r1}
%\definechangesauthor[name=2, color=teal]{r2}
%\definechangesauthor[name=3, color=orange]{r3}
%\definechangesauthor[name=4, color=violet]{ed}

\begin{document}

\noindent \today\\[2em]
Marios Karouzos, Senior Editor\\
Nature Astronomy\\

\bigskip

\noindent Dear Dr. Karouzos, \\

Please accept our heartfelt thanks for allowing us to resubmit this manuscript after such a long delay.  This past year has been challenging in many ways, and we are deeply appreciative that you have held the door open for us this long.  We are all still very excited by the discovery presented here, and feel it will be well received in the community.

We provide two updated versions of the manuscript.  The ``changetext'' version includes line numbers and colored text, showing changes since the original submission.  The ``revised'' version is just the current manuscript text, without showing changes.  Our estimate of the relevant sizes in the revised document are:

\begin{itemize}
    \item Main Text : 2009 words.
    \item First Paragraph : 204 words.
    \item Main text contains 3 figures and 1 table.
    \item Figure 1 legend : 195 words
    \item Figure 2 legend : 283 words
    \item Figure 3 legend : 122 words
    \item Table 1 legend : 30 words
    \item Methods Text :  4124 words.
    \item The Methods section contains 6 tables and 10 figures.
    \item Figure and Table legends in Methods and Supplementary Material are all $<350$ words.  
    \item Figures in the manuscript are shown at their estimated final size.
    \item Total size of the revised main text document is 2.5 MB.
    \item Total size of the Methods+Supplementary+References is 8.5 MB
\end{itemize}

Our point-by-point response to the referees is provided as a separate document.  There are 4 versions: one for each referee with their comments alone, and one that includes all comments together. In addition, we have made adjustments for style, following your guidance.  Where appropriate, these changes are highlighted with purple text in the changetext manuscript.  We have:

\begin{itemize}
\item{re-worded the first paragraph to follow the guidelines you provided.  We believe it now conforms to the Nature Astronomy style, and it is now 200 words according to our count;}
\item{%You encouraged us to archive the data for this manuscript in an accessible, persistent repository. We have 
prepared a github repository (currently private) with all of the code used to construct the key figures in the paper, along with data tables recording the SN photometry.  We will change  this repository to public upon acceptance of the manuscript and include a link in the final published article;}
\item{adjusted figure legends to ensure that they provide a brief description of the figure and the symbols used, within 350 words, including definitions of any error bars employed in the figures;}
\item{adjusted the reference list to keep it at 30 references in the main text;}
\item{added a statement about the corresponding author for data requests, and ensured that we have a statement of author contributions.}
\end{itemize}

Please express our sincere gratitude to the referees for their helpful comments and extraordinary patience.

\bigskip

\noindent Warm Regards,\\
\smallskip
Steve Rodney, on behalf of co-authors

\end{document}

### 2021-05-21 Paper/NatureAstronomy/main_resubmission1.tex

\usepackage[utf8]{inputenc}
%\usepackage[backend=biber,style=nature,citestyle=numeric]{biblatex}
%\usepackage{biblatex}
%\addbibresource{zotero_references_biblatex.bib}

%extra packages added during drafting, remove for submission
\usepackage{hyperref}
\usepackage{booktabs}

%\usepackage{amssymb}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage[draft,xcolor=usenames]{changes}
\definechangesauthor[name=1, color=blue]{r1}
\definechangesauthor[name=2, color=teal]{r2}
\definechangesauthor[name=3, color=orange]{r3}
\setauthormarkup{}

\gdef\arcsec{$^{\prime\prime}$}
\def\TNSname{AT 2016tbd\xspace}
\def\SNABC{AT 2016tbd\xspace}
%\def\SNABC{SN Requiem\xspace}
%\def\reqlike{Requiem-like\xspace}
\def\reqlike{16tbd-like\xspace}

@@

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% http://www.sciencemag.org/authors/preparing-manuscripts-using-latex 
% This package should properly format in-text
% reference calls and reference-list numbers.

%\usepackage{scicite}
%\usepackage{nature}

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

@@

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% Include your paper's title here

\title{A Gravitationally Lensed Supernova with an Observable Two-Decade Time Delay}

@@

\author{
 Steven A. Rodney$^{1\ast}$, 
 Gabriel Brammer$^{2\ast}$,  
 Justin D.\,R.~Pierel$^{1}$,\\ 
 Johan Richard$^{3}$, 
 Sune Toft$^{2}$,\\  
 Kyle F. O'Connor$^{1}$, 
 Mohammad Akhshik$^{4}$, 
 Katherine E. Whitaker$^{2,4,5}$  
\\
\normalsize{$^{1}$Department of Physics \& Astronomy, University of South Carolina,}\\ \normalsize{ Columbia, SC 29208, USA}\\
\normalsize{$^{2}$Cosmic Dawn Center, Niels Bohr Institute, University of Copenhagen,}\\ \normalsize{Lyngbyvej 2, Copenhagen \O\ 2100 Denmark}\\
\normalsize{$^{3}$Univ Lyon, Univ Lyon1, Ens de Lyon, CNRS, Centre de Recherche Astrophysique de Lyon}\\ \normalsize{UMR5574, F-69230, Saint-Genis-Laval, France}\\
\normalsize{$^{4}$Department of Physics, University of Connecticut,}\\ \normalsize{Storrs, CT 06269, USA}\\
\normalsize{$^{5}$Department of Astronomy, University of Massachusetts,}\\ \normalsize{Amherst, MA 01003, USA}\\
\normalsize{$^\ast$To whom correspondence should be addressed;}\\ \normalsize{E-mail:  srodney@sc.edu, gabriel.brammer@nbi.ku.dk}
}

@@

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}

  We report the discovery of a strongly-lensed supernova (SN) explosion that will enable a time delay measurement with an uncertainty of $<1\%$. This unprecedented scenario will enable a rare and highly precise measurement of the Hubble\added[id=r1]{-LeMa\^{i}tre} constant from a single lensed SN.  The SN is in an evolved galaxy at $z=1.95$, gravitationally lensed by the massive foreground galaxy cluster MACS J0138.0-2155.  This is the most distant multiply-imaged supernova yet discovered.  We classify the supernova as a Type Ia with {$>90\%$} probability based on archival imaging observations from the \textit{Hubble Space Telescope}.  Three lensed images of the SN are detected with relative time delays of $<$200 days, and we predict that a fourth image in the component closest to the cluster core will appear in the year 2037$\pm$2.  Observing the light curve of that future image could provide a time delay precision of $\approx 7$ days over an extraordinary 20 year baseline.

\end{sciabstract}

@@

Observations of Type Ia supernova (SN) explosions have played a key role in mapping the cosmic expansion history, and led to the discovery of dark energy that now appears to be driving an accelerating cosmic expansion rate 
\cite{riess_observational_1998,perlmutter_measurements_1999, Riess_large_2019}.  Determining the nature of dark energy and how it may evolve over time is a primary goal for the large-scale cosmology experiments of the 2020s
\cite{amendola_cosmology_2013,spergel_wide_2015,Ivezic_lsst_2019}.
A series of recent investigations into the expansion rate of the universe (the Hubble-LeMa\^itre constant; $H_0$) have found that measurements from the local universe are significantly different from the value inferred from measurement of the cosmic microwave background (CMB) radiation \cite{Riess_large_2019,aghanim_planck_2018}.  Resolving this tension in $H_0$ measurements may reveal new physics from the early universe, and to do so requires multiple 
independent cosmological probes \cite{verde_tensions_2019}.

One promising cosmological tool uses %time delays of 
gravitational lens systems in which a background
source appears as multiple images that arrive to the observer with relative delays \cite{einstein_lens_1936}.
%, which %for a given lens potential 
When such a strongly-lensed source is variable, one can measure the time delay between any pair of lensed images and derive a ratio of angular diameter distances to the foreground lens and background source. 
This distance ratio is sensitive to cosmological parameters, such as 
%the expansion rate of the universe, 
$H_0$  and the dark energy equation-of-state, $w$ \cite{refsdal_possibility_1964,coe_cosmological_2009,linder_lensing_2011}.
In recent years this method has been successfully applied to lensed
quasars \cite{suyu_cosmology_2014,bonvin_cosmograil_2019}, with six high-precision measurements to date \cite{wong_h0licow_2019}. As this sample of time delay lenses grows to several dozen, it is expected to deliver a measurement of $H_0$ with 1\% precision \cite{suyu_cosmological_2018}. \added[id=r1]{A future sample with $\sim100$ well-measured lensing time delays will be a competitive tool for dark energy studies \cite{linder_lensing_2011,jee_time-delay_2016,suyu_HOLISMOKES_2020}.}

Gravitationally lensed SN with multiple images present an attractive new addition to this field. Most notably, SN exhibit relatively simple photometric behavior, with well-understood light curve shapes and colors---in contrast to the stochastic variation of quasars.  However, despite the promise of lensed SN time delay cosmography as a cosmological distance probe, to date, there have been only two lensed SN observed with multiple images \cite{kelly_multiple_2015,goobar_iptf16geu:_2017}. The first, {\it SN Refsdal}, was a peculiar Type II SN whose image with the longest delay was missed \cite{kelly_sn_2016}. The second, {\it SN 2016geu}, was a Type Ia SN with short delays that make high precision time delay measurements impossible \cite{dhawan_magnification_2019}. \added[id=r1]{An earlier discovery of an unusually luminous SN \cite{chornock_ps1-10afx_2013} was also shown to be a strongly-lensed SNIa, 
\cite{quimby_extraordinary_2013,quimby_detection_2014}
though the multiple images were not resolved.}  Here we present the discovery 
of a third lensed SN resolved into multiple images: \SNABC.

We discovered \SNABC ({\it AstroNotes citation})
using data from the Hubble Space Telescope (HST) program {\it REsolved QUIEscent Magnified Galaxies} (REQUIEM, HST-GO-15663, PI:Akhshik). This project targets rare examples of massive galaxies with low specific star formation rates that have been magnified by strong gravitational lensing. %These are typically too faint and compact to be resolved even with Hubble, but due to lensing are magnified. This provides a unique opportunity to  understand the still mysterious quenching mechnism. 
The brightest and most spectacular galaxy targeted by REQUIEM
is MRG-M0138, first discovered as a massive red galaxy (MRG) at $z=1.95$ \cite{newman_resolving_2018} behind the massive galaxy cluster MACS J0138.0-2155 \cite{ebeling_macs_2001}.
MRG-M0138 is quadruply lensed by a foreground galaxy cluster at $z=0.338$.  
During analysis of REQUIEM observations obtained 13--14 July 2019 \cite{materials_methods_2020} we discovered three point 
sources that were present in archival HST images from 18--19 July 2016, part of the program 
that first confirmed the MRG-M0138 galaxy as a strongly-lensed object (HST-GO-14496; PI:Newman). 
Each point source is within 5 arcseconds of one of the four MRG-M0138 images.  None of the
three point sources are present in the REQUIEM HST data in 2019 (Fig. \ref{fig:layout}). We infer that 
the point sources are multiple images of a single astrophysical 
transient in MRG-M0138, most likely a SN \cite{materials_methods_2020}.  
%The official designation for this multiply-imaged astrophysical transient is \TNSname. 
%Formal designation as a SN requires spectroscopic confirmation, but with a high-confidence photometric classification we coin the name \SNABC for this unusual source.

To construct a lens model for the MACS J0138.0-2155 cluster we use the \lenstool software \cite{jullo_bayesian_2007,kneib_lenstool_2011} to model the mass distribution in the cluster core as the combination of a cluster-scale and multiple galaxy-scale potentials \cite{materials_methods_2020}.  To avoid unintended bias, we kept the lens model development completely separate from the analysis of the SN.  Only upon independent completion of both were the results combined for the analysis described here.  The input model constraints are the positions and redshifts of the MRG-M0138 galaxy at $z=1.95$ (both the central light peak and the SN location in each image) as well as a multiply-imaged background 
galaxy at $z=0.766$, both having secure spectroscopic redshifts \cite{materials_methods_2020}.  
From this model we derive estimates for the lensing magnification and time delay of each of the SN images, including the predicted fourth image (Table \ref{tab:time_delays}).
The lensing model predicts that the SN should appear in the fourth MRG-M0138 image in the year 2037$\pm$1.5, demagnified with $\mu=0.4\pm0.2$. The predicted location for this future image is $\alpha=01:38:04.15,\delta=-21:55:24.659$.
A fifth image will also appear at date later still, located near the center of the cluster and much more significantly demagnified, so it will not be easily observable.

To realize the cosmological promise of \SNABC 
we need to estimate the age of each SN image, which in turn constrains the lensing time delays. For this
goal, it is necessary to have a firm determination of the transient’s class. Expected time delays and magnifications from the lens model exclude any of the various rapidly evolving and low-luminosity stellar transient classes, strongly suggesting that it is a SN. The first-order SN distinction remaining is between a Type Ia SN---the explosion of a white dwarf star in a binary system---and a core collapse SN (CCSN)---the end-point of a star with mass $>10 M_{\odot}$. 
The properties of the host galaxy can inform this classification because CCSN are limited to galaxies with young stellar populations. Limits on the specific star formation rate and age for this host, MRG-M0138, show it to be a massive but very quiescent and evolved galaxy, %($\log(M_*/M_\odot)=11.7\pm0.2; )$ , 
unlikely to retain any significant population of high-mass stars \cite{newman_resolving_2018}. Based on observed properties of the host galaxy alone, we find a 62-75\% probability that \SNABC is a Type Ia SN \cite{materials_methods_2020}. Adopting the lens model magnifications for the three observed SN images (Table \ref{tab:time_delays}) we can locate each SN image in color-magnitude space (Fig. \ref{fig:class}). This provides more classification leverage, yielding p(Ia)=95\% \cite{materials_methods_2020}. By also including the model-predicted time delays, we can treat the three SN images effectively as three points on a common SN light curve, and we find p(Ia)=94\% \cite{materials_methods_2020}.
An improved classification can be achieved with spectroscopy and multi-band photometry upon arrival of the fourth image.

The color of a SNIa evolves substantially over its lifetime as the photosphere expands and cools, revealing different layers of the expanding shell and driving episodes of recombination \cite{woosley_type_2007,kasen_type_2009}. Since the phenomenon of gravitational lensing in general is achromatic, this color evolution makes it possible to derive an age constraint that is largely independent of the lens model \cite{materials_methods_2020}.  Combining this information with magnification constraints from the lens modeling helps break parameter degeneracies, yielding the measured delays in Table \ref{tab:time_delays} \cite{materials_methods_2020}. 
Remarkably, the ages of image 1 and 3 are constrained to better than $\pm$20 days, despite having only a single epoch of photometric data.    These uncertainties may be 
further reduced when the future fourth image is observed with high-precision, multi-epoch photometry.  Such a light curve will pin down the intrinsic SN light curve parameters that are shared by all images, and break remaining parameter degeneracies (see supplemental materials).

\added[id=r1]{By itself, \SNABC will not be particularly powerful as a solitary cosmological probe.  However,} future large-scale surveys such as the Vera C. Rubin Observatory 
and the Nancy Grace Roman Space Telescope will observe dozens to hundreds of lensed SN over their mission lifetimes, \added[id=r1]{though the vast majority of these will be lensed by galaxy-scale deflectors}  \cite{oguri_gravitationally_2010,goldstein_rates_2019,wojtak_magnified_2019}. \deleted[id=r1]{However, many of these lensed SN will require substantial follow-up resources to be made cosmologically useful.} Thus, almost all will have significantly shorter 
delays (of order 10--100 days) and therefore higher relative uncertainties than \SNABC.  Since it is the {\it fractional} time delay uncertainty that 
propagates through to any time delay distance measurement, the extraordinarily long time delays of cluster-lensed SNe like \SNABC can deliver significantly better time delay precision, with comparable observational cost.  In fact, the long time baseline of lensed SN like \SNABC effectively insures that their cosmological precision is not limited by time delay measurement uncertainty.

@@

To achieve these cosmological constraints in practice, such a sample could be developed with regular monitoring of cluster-scale lenses, partnered with modest follow-up to characterize any lensed SN discovered.  As \SNABC shows, even a single epoch of photometric data at the time of detection may be sufficient to provide the necessary anchor point for this promising cosmological tool.  %After that, all we need is patience.

@@

\midrule
\textit{1}  & $87^{+16}_{-13}$ & -- & $5\pm1$ & --\\
\textit{2} & $-20^{+20}_{-11}$ & $106^{+25}_{-24}$ & $7\pm3$ & $82\pm62$ \\
\textit{3} & $100^{+27}_{-31}$ & $-10\pm24$ &$3.9\pm0.5$  & $-19\pm34$ \\
\textit{4} & -- & -- & $0.4\pm0.2$ & $7742\pm540$\\

@@

\clearpage
%\section*{Supplementary materials}
%Materials and Methods\\
%Supplementary Text\\
%Figures S1 to S8\\
%Tables S1 to S7\\
%References \textit{(32-68)}
%\clearpage
%\setcounter{table}{0}
%\renewcommand{\thetable}{S\arabic{table}}

@@

%\clearpage

@@

\subsection*{Observations of the SN and Lensing System}

@@

\subsection*{VLT Spectroscopy}
\label{sec:vltmuse}

@@

\subsection*{SN Sub-Classification Based on SN Photometry}

>>>>

\usepackage[utf8]{inputenc}

% Required packages
\usepackage{caption}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{graphicx}

%extra packages added during drafting, remove for submission
\usepackage{lineno}

%%%  SELECT DRAFT VERSION (CHANGETEXT) OR FINAL 
%\usepackage[draft,xcolor=dvipsnames]{changes}
%\linenumbers
\usepackage[final,xcolor=dvipsnames]{changes}

\definechangesauthor[name=1, color=blue]{r1}
\definechangesauthor[name=2, color=teal]{r2}
\definechangesauthor[name=3, color=Maroon]{r3}
\definechangesauthor[name=4, color=violet]{ed}
\setauthormarkup{}

\gdef\arcsec{$^{\prime\prime}$}
\def\TNSname{AT 2016tbd\xspace}
\def\SNABC{AT 2016tbd\xspace}
\def\reqlike{16tbd-like\xspace}

@@

%\usepackage{nature}
\usepackage{times}

@@

%The next command sets up an environment for the abstract to your paper.
\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% Include your paper's title here
\title{A Gravitationally Lensed Supernova with an Observable Two-Decade Time Delay}

@@

\author{
 Steven A. Rodney$^{1\ast}$, 
 Gabriel Brammer$^{2,3\ast}$,  
 Justin D.\,R.~Pierel$^{1}$,\\ 
 Johan Richard$^{4}$, 
 Sune Toft$^{2,3}$,\\  
 Kyle F. O'Connor$^{1}$, 
 Mohammad Akhshik$^{5}$, 
 Katherine E. Whitaker$^{6,2}$  
\\

\normalsize{$^{1}$Department of Physics \& Astronomy, University of South Carolina,}\\ \normalsize{ Columbia, SC 29208, USA}\\
\normalsize{$^{2}$Cosmic Dawn Center (DAWN), Copenhagen Denmark}\\
\normalsize{$^{3}$Niels Bohr Institute, University of Copenhagen, Jagtvej 128, Copenhagen, Denmark}\\
\normalsize{$^{4}$Univ Lyon, Univ Lyon1, Ens de Lyon, CNRS, Centre de Recherche Astrophysique de Lyon}\\ \normalsize{UMR5574, F-69230, Saint-Genis-Laval, France}\\
\normalsize{$^{5}$Department of Physics, University of Connecticut,}\\ \normalsize{Storrs, CT 06269, USA}\\
\normalsize{$^{6}$Department of Astronomy, University of Massachusetts,}\\ \normalsize{Amherst, MA 01003, USA}\\
\normalsize{$^\ast$To whom correspondence should be addressed;}\\ \normalsize{E-mail:  srodney@sc.edu, gabriel.brammer@nbi.ku.dk}
}

@@

% Place your abstract within the special {sciabstract} environment.
\clearpage
\begin{sciabstract}

\replaced[id=ed]{
When the light from a distant stellar explosion passes very near to a foreground galaxy or cluster, gravitational lensing can cause it to appear as multiple images on the sky.  Such strongly-lensed supernovae can be used to constrain the cosmic expansion rate \cite{refsdal_possibility_1964} and dark energy models \cite{treu_time_2016}.  Achieving these cosmological goals requires many more lensed supernova discoveries and more efficient observational follow-up.   Here we report the discovery of a multiply-imaged supernova that will enable a time delay measurement with an uncertainty of $<1\%$.  It appeared in an evolved galaxy at $z=1.95$, gravitationally lensed by a massive foreground galaxy cluster. It is likely a Type Ia supernova---the explosion of a low-mass stellar remnant, whose light curve can be used to measure cosmic distances.  In archival Hubble Space Telescope imaging, three lensed images of the supernova are detected with relative time delays of $<$200 days.  We predict a fourth image will appear close to the cluster core in the year 2037$\pm$2.   Observation and modeling of that future image could provide a time delay precision of $\approx 7$ days over an extraordinary 20 year baseline.  This object demonstrates that cosmologically useful time delays can be measured with minimal observational cost using cluster-lensed supernovae.}{
  We report the discovery of a strongly-lensed supernova (SN) explosion that will enable a time delay measurement with an uncertainty of $<1\%$. This unprecedented scenario will enable a rare and highly precise measurement of the Hubble\added[id=r1]{-LeMa\^{i}tre} constant from a single lensed SN.  The SN is in an evolved galaxy at $z=1.95$, gravitationally lensed by the massive foreground galaxy cluster MACS J0138.0-2155.  This is the most distant multiply-imaged supernova yet discovered.  We classify the supernova as \added[id=r2]{most likely} a Type Ia \deleted[id=r2]{with {$>90\%$} probability} based on archival imaging observations from the \textit{Hubble Space Telescope}.  Three lensed images of the SN are detected with relative time delays of $<$200 days, and we predict that a fourth image in the component closest to the cluster core will appear in the year 2037$\pm$2.  Observing the light curve of that future image could provide a time delay precision of $\approx 7$ days over an extraordinary 20 year baseline. 
}
\end{sciabstract}

@@

Observations of Type Ia supernova (SN) explosions have played a key role in mapping the cosmic expansion history, and led to the discovery of dark energy that now appears to be driving an accelerating cosmic expansion rate 
\cite{riess_observational_1998,perlmutter_measurements_1999} %Riess_large_2019}.  
Determining the nature of dark energy and how it may evolve over time is a primary goal for the large-scale cosmology experiments of the 2020s
\cite{amendola_cosmology_2013,spergel_wide_2015,Ivezic_lsst_2019}.
\deleted[id=ed]{A series of} Recent investigations into the expansion rate of the universe (the Hubble-LeMa\^itre constant; $H_0$) have found that measurements from the local universe are significantly different from the value inferred from measurement of the cosmic microwave background \deleted[id=ed]{(CMB)} radiation \cite{Riess_large_2019,aghanim_planck_2018}.  Resolving this tension in $H_0$ measurements may reveal new physics from the early universe, and to do so requires multiple 
independent cosmological probes \cite{verde_tensions_2019}.

One promising cosmological tool uses %time delays of 
gravitational lens systems in which a background
source appears as multiple images that arrive to the observer with relative delays \cite{einstein_lens_1936}.
%, which %for a given lens potential 
When such a strongly-lensed source is variable, one can measure the time delay between any pair of lensed images and derive a ratio of angular diameter distances to the foreground lens and background source. 
This distance ratio is sensitive to cosmological parameters, such as 
%the expansion rate of the universe, 
$H_0$ \cite{refsdal_possibility_1964} and the dark energy equation-of-state, $w$ \cite{coe_cosmological_2009}.%,linder_lensing_2011}.
In recent years this method has been successfully applied to lensed
quasars %cite{suyu_cosmology_2014,bonvin_cosmograil_2019}, 
with six high-precision measurements to date \cite{wong_h0licow_2019}. As this sample of time delay lenses grows to several dozen, it is expected to deliver a measurement of $H_0$ with 1\% precision \cite{suyu_cosmological_2018}. \added[id=r1]{A future sample with $\sim100$ well-measured lensing time delays will be a competitive tool for dark energy studies \cite{treu_time_2016}.}
%\cite{linder_lensing_2011,jee_time-delay_2016,suyu_HOLISMOKES_2020}.}

Gravitationally lensed SN with multiple images present an attractive new addition to this field. Most notably, SN exhibit relatively simple photometric behavior, with well-understood light curve shapes and colors---in contrast to the stochastic variation of quasars.  However, \deleted[id=ed]{despite the promise of lensed SN time delay cosmography as a cosmological distance probe,} to date there have been only two lensed SN observed with multiple images \cite{kelly_multiple_2015,goobar_iptf16geu:_2017}. The first, {\it SN Refsdal}, was a peculiar Type II SN whose image with the longest delay was missed \cite{kelly_sn_2016}. The second, {\it SN 2016geu}, was a Type Ia SN with short delays that make high precision time delay measurements impossible \cite{dhawan_magnification_2019}. \added[id=r1]{An earlier discovery of an unusually luminous SN \cite{chornock_ps1-10afx_2013} was also shown to be a strongly-lensed SNIa,  \cite{quimby_detection_2014} %quimby_extraordinary_2013,
though the multiple images were not resolved.}  Here we present the discovery 
of a third lensed SN resolved into multiple images: \SNABC.

We discovered \SNABC ({\it AstroNotes citation})
using data from the Hubble Space Telescope (HST) program {\it REsolved QUIEscent Magnified Galaxies} (REQUIEM, HST-GO-15663, PI:Akhshik). This project targets rare examples of massive galaxies with low specific star formation rates that have been magnified by strong gravitational lensing. %These are typically too faint and compact to be resolved even with Hubble, but due to lensing are magnified. This provides a unique opportunity to  understand the still mysterious quenching mechnism. 
The brightest and most spectacular galaxy targeted by REQUIEM
is MRG-M0138, first discovered as a massive red galaxy (MRG) at $z=1.95$ \cite{newman_resolving_2018} behind the massive galaxy cluster MACS J0138.0-2155 \cite{ebeling_macs_2001}.
MRG-M0138 is quadruply lensed by a foreground galaxy cluster at $z=0.338$.  
During analysis of REQUIEM observations obtained 13--14 July 2019
%\cite{materials_methods_2020} 
we discovered three point 
sources that were present in archival HST images from 18--19 July 2016, part of the program (HST-GO-14496; PI:Newman)
that first confirmed the MRG-M0138 galaxy as a strongly-lensed object 
\added[id=ed]{(Methods: Observations)}. 
Each point source is within 5 arcseconds of one of the four MRG-M0138 images.  None of the
three point sources are present in the REQUIEM HST data in 2019 (Fig. \ref{fig:layout}). We infer that 
the point sources are multiple images of a single astrophysical 
transient in MRG-M0138, most likely a SN. %\cite{materials_methods_2020}.  
%The official designation for this multiply-imaged astrophysical transient is \TNSname. 
%Formal designation as a SN requires spectroscopic confirmation, but with a high-confidence photometric classification we coin the name \SNABC for this unusual source.

To construct a lens model for the MACS J0138.0-2155 cluster we use the \lenstool software \cite{jullo_bayesian_2007,kneib_lenstool_2011} to model the mass distribution in the cluster core as the combination of a cluster-scale and multiple galaxy-scale potentials 
\added[id=ed]{(Methods: Lens Modeling)}.
%\cite{materials_methods_2020}.  
To avoid unintended bias, we kept the lens model development completely separate from the analysis of the SN.  Only upon \deleted[id=ed]{independent} completion of both were the results combined for the analysis described here.  The input model constraints are the positions and redshifts of the MRG-M0138 galaxy at $z=1.95$ (both the \replaced[id=r3]{galaxy's centroid position }{central light peak} and the SN location in each image) as well as a multiply-imaged background 
galaxy at $z=0.766$, both having secure spectroscopic redshifts
\added[id=ed]{(Methods: Observations)}.
%\cite{materials_methods_2020}.  
From this model we derive estimates for the lensing magnification and time delay of each of the SN images, including the predicted fourth image (Table \ref{tab:time_delays}).
The lensing model predicts that the SN should appear in the fourth MRG-M0138 image in the year 2037$\pm$\replaced[id=r1]{2}{1.5}, demagnified with $\mu=0.4\pm0.2$. The predicted location for this future image is $\alpha=$01:38:04.15, $\delta=-$21:55:24.659.
A fifth image will also appear at a still later date, located near the center of the cluster and much more significantly demagnified, so it will not be easily observable.  \added[id=r1]{We anticipate that future lens modeling of the cluster may improve on these predictions} \added[id=ed]{(Supplementary Note: Future Work}).

To realize the cosmological promise of \SNABC 
we need to estimate the age of each SN image, which in turn constrains the lensing time delays. For this
goal, it is helpful to have a firm determination of the transient’s class. Expected time delays and magnifications from the lens model exclude any of the various rapidly evolving and low-luminosity stellar transient classes, strongly suggesting that it is a SN. The first-order SN distinction remaining is between a Type Ia SN---the explosion of a white dwarf star in a binary system---and a core collapse SN (CCSN)---the end-point of a star with mass $>10 M_{\odot}$. 
The properties of the host galaxy can inform this classification because CCSN are limited to galaxies with young stellar populations. Limits on the specific star formation rate and age for this host, MRG-M0138, show it to be a massive but very quiescent and evolved galaxy, %($\log(M_*/M_\odot)=11.7\pm0.2; )$ , 
unlikely to retain any significant population of high-mass stars \cite{newman_resolving_2018}. Based on observed properties of the host galaxy alone, we find a 62-75\% probability that \SNABC is a Type Ia SN. Adopting the lens model magnifications for the three observed SN images (Table \ref{tab:time_delays}) we can locate each SN image in color-magnitude space (Fig. \ref{fig:class}). \replaced[id=r1]{After magnification correction, all three images are still brighter than expected for a Type Ia SN, which may indicate that a lens-modeling degeneracy is at play \cite{falco_model-dependent_1985, schneider_source-position_2014}. Nevertheless, the magnification-corrected \SNABC data are more consistent with the Type Ia population than any CC SN sub-class (Fig.~\ref{fig:class}a,b), and demonstrate the expected evolution of a Type Ia SN color and brightness over $\sim 100$ days (Fig.~\ref{fig:class}c).}{This provides more classification leverage, yielding p(Ia)=95\%.}
By also including the model-predicted time delays, we can treat the three SN images effectively as three points on a common SN light curve, and we find p(Ia)=94\% 
\added[id=ed]{(Methods: Classification)}.
%\cite{materials_methods_2020}. 
\replaced[id=r2]{An improved classification can be achieved }{These final classifications are strongly conditional on the priors, and we look forward to an improved classification} with spectroscopy and multi-band photometry upon arrival of the fourth image.  \added[id=r2]{For the remainder of this analysis, we proceed under the assumption that \SNABC is indeed a Type Ia SN---though the subsequent analysis could be revised to achieve similar results with a different underlying SN model.}

The color of a SNIa evolves substantially over its lifetime as the photosphere expands and cools, revealing different layers of the expanding shell and driving episodes of recombination \cite{kasen_type_2009}. %woosley_type_2007
Since the phenomenon of gravitational lensing in general is achromatic, this color evolution makes it possible to derive an age constraint that is largely independent of the lens model. %\cite{materials_methods_2020}.  
Combining this information with magnification constraints from the lens modeling helps break parameter degeneracies, yielding the measured delays in Table \ref{tab:time_delays}
\added[id=ed]{(Methods: Time Delay)}.
%\cite{materials_methods_2020}. 
Remarkably, the ages of image 1 and 3 are constrained to better than $\pm$20 days, despite having only a single epoch of photometric data.    These uncertainties may be 
further reduced when the future fourth image is observed with high-precision, multi-epoch photometry.  Such a light curve will pin down the intrinsic SN light curve parameters that are shared by all images, and break remaining parameter degeneracies.
%\added[id=ed]{(Supplementary Note: Future Observations)}.

\added[id=r1]{As with any single time-delay system, \SNABC will not be particularly powerful as a solitary cosmological probe.  However,} future large-scale surveys such as the Vera C. Rubin Observatory 
and the Nancy Grace Roman Space Telescope will observe dozens to hundreds of lensed SN over their mission lifetimes, \added[id=r1]{though the vast majority of these will be lensed by galaxy-scale deflectors \cite{oguri_gravitationally_2010,  goldstein_rates_2019, pierel_projected_2020}.} %wojtak_magnified_2019, 
\deleted[id=r1]{However, many of these lensed SN will require substantial follow-up resources to be made cosmologically useful.} Thus, almost all will have significantly shorter 
delays (of order 10--100 days) and therefore higher relative uncertainties than \SNABC.  Since it is the {\it fractional} time delay uncertainty that 
propagates through to any time delay distance measurement, the extraordinarily long time delays of cluster-lensed SNe like \SNABC can deliver significantly better time delay precision, with comparable observational cost.  In fact, the long time baseline of lensed SN like \SNABC effectively insures that their cosmological precision is not limited by time delay measurement uncertainty. 
\added[id=r1]{Cluster-scale lenses are more complex than galaxy-scale lenses, but they generally have ``independent'' measurements of the magnification from several multiply-imaged systems in the same field. As such, they can provide a valuable check on systematics related to specific lens modeling software, microlensing variation, and more.}

@@

To achieve these cosmological constraints in practice, such a sample could be developed with regular monitoring of cluster-scale lenses, partnered with modest follow-up to characterize any lensed SN discovered
\added[id=ed]{(Supplementary Note: Future Work}).
%\cite{materials_methods_2020}.  
As \SNABC shows, even a single epoch of photometric data at the time of detection may be sufficient to provide the necessary anchor point for this promising cosmological tool.  
%After that, all we need is patience.

@@

\midrule
\textit{1}  & $92^{+21}_{-19}$ & -- & $5\pm1$ & --\\
\textit{2} & $-24^{+16}_{-7}$ & $114^{+28}_{-31}$ & $7\pm3$ & $82\pm62$ \\
\textit{3} & $107^{+26}_{-21}$ & $-17^{+19}_{-16}$ &$3.9\pm0.5$  & $-19\pm34$ \\
\textit{4} & -- & -- & $0.4\pm0.2$ & $7742\pm540$\\
\textit{5} & -- & -- & $<0.01$ & $9463\pm770$\\

@@

%\section*{Supplementary materials}
%Materials and Methods\\
%Supplementary Text\\
%Figures S1 to S8\\
%Tables S1 to S7\\
%References \textit{(32-68)}
%\clearpage
%\setcounter{table}{0}
%\renewcommand{\thetable}{S\arabic{table}}

@@

%\clearpage
%\end{document}

@@

\subsection*{Observations} % of the SN and Lensing System}

@@

\subsubsection*{VLT Spectroscopy}
\label{sec:vltmuse}

@@

\subsubsection*{SN Sub-Classification Based on SN Photometry}
